### 2018-01-18 usenix18/ld.tex
Usenix template

\subsubsection*{Source language} Our source language
 is a simple imperative language shown in \figureref{srclang}. Types
$\psi$ consist of the base types $\sigma$, and arrays
of base types $\sigma[n]$, where $n$ is the array length. Though we
model only one dimensional arrays, our implementation supports
 multi-dimensional arrays as well. Expressions $e$ in the language include the
 integer constants $n$, \kw{bool} constants $\top$ and $\bot$,
variables $x$, binary operations $e_{1} \mulop e_{2}$ and $e_{1} >
e_{2}$ (we support several other
operators in the implementation, detailed in Section~\ref{sec:impl}),
conditionals $\cond{e}{e_{1}}{e_{2}}$,
array literals
$[\overline{e_{i}}]_{n}$\footnote{We write $\overline{e}$ (and
  similarly for other symbols) to denote a sequence of expressions.
The length of the sequence is usually clear from the context.}, and
array reads $x[e]$. The expression $\kw{in}_{j}$ denotes input from
party $j$. The statements $s$ in the language comprise of variable
declarations and assignments ($\psi\:x = e$ and $x := e$ resp.),
\kw{for} loops, array writes ($x[e_{1}] := e_{2}$), \kw{if}
statements, and sequence of statements ($s_{1}; s_{2}$). The statement
$\kw{out}\:e$ denotes revealing the value of $e$ to the
parties. The \kw{while} statement is an internal syntax that is not
exposed to the programmer. \\

@@

The judgment $\rho \vdash e \Downarrow v$ denotes the big-step
evaluation of an expression~$e$ to a value~$v$ under the runtime
environment~$\rho$. Rule ({\sc{E-Var}}) looks up the value of $x$ in
the environment. Rule ({\sc{E-Mult}}) inductively evaluates $e_{1}$ and
$e_{2}$, and returns their product. Rule ({\sc{E-Read}})
evaluates an array read operation. It first evaluates $x$ to an array
value $[\overline{c_{i}}]_{n_{1}}$, and $e$ to a \kw{uint} value
$n$. It then returns $c_{n}$, the $n$-th index value in the array,
provided $n < n_{1}$, the length of the array. Rule ({\sc{E-Inp}})
evaluates to some constant $c$ denoting party $j$'s input. 
%We model the inputs to be base constants, a
An array input can be written in the
language as $[\kw{in}_{j}]_{n}$, which can then evaluate using the
rule ({\sc{E-Arr}}). The notation $\forall i \in [n]$ is read
  as $\forall i \in \{0\dots n - 1\}$. The remaining rules are
straightforward, and are elided for space reasons.

@@

The judgment $\Gamma \vdash e : \tau \leadsto \widetilde{e}$, where
$\Gamma$ maps variables $x$ to types $\tau$, says that
under $\Gamma$, $e$ (in the source language) compiles to $\widetilde{e}$
(in the intermediate language), where $\widetilde{e}$ has type $\tau$.
Rules ({\sc{T-UInt}}) and ({\sc{T-Bool}}) assigns the label
$\mathcal{P}$ to the constants, as the constants are always
public. Rule ({\sc{T-Mult}})
compiles a multiplication to either a public multiplication  ($\mulop_{\mathcal{P}}$), or a
secret arithmetic multiplication ($\mulop_{\mathcal{A}}$). As
our compiler is cryptographic cost aware, it never compiles the
multiplication to boolean multiplication $\mulop_{\mathcal{B}}$ (\sectionref{impl}). In a similar manner,
rule ({\sc{T-Gt}}) compiles $e_{1} > e_{2}$ to either public comparison,
or secret boolean comparison $>_\mathcal{B}$ (and never
$>_{\mathcal{A}}$).
%
The rule for conditional
({\sc{T-Cond}}) has two cases: when the conditional expression $e$ is of
type $\kw{bool}^{\mathcal{P}}$, both the branches have a base type
$\sigma^{\ell_{1}}$, for an arbitrary $\ell_{1}$, and the conditional is
compiled to a public conditional, whereas when the conditional
expression has type $\kw{bool}^{\mathcal{B}}$, $\ell_{1}$ is also
$\mathcal{B}$, and the conditional is compiled to a secret conditional
using a boolean circuit. Note that we restrict the type of the branches
to be of base type. 
Rule ({\sc{T-Read}}) type checks an array read. It checks
that the array index $e$ is public, and uses a static bounds checking
judgment $ \models e < n$ to prove that the array index is in
bounds. Section~\ref{sec:impl} discusses our implementation of this
check. Rule ({\sc{T-Inp}}) picks a label $m$ for the input. Finally,
the rule ({\sc{T-Sub}}) is the subsumption rule that coerces an expression of
type $\sigma^{\ell}$ to an expression of type $\sigma^{m}$ using the
coerce expression. It is important for security that the secrets
cannot be coerced to public values and indeed ({\sc{T-Sub}}) does not
permit it.

@@

We use $\widehat\rho_j$ to denote the circuit environment for party $j.$ 
We give the 
judgments $\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \gate
\Downarrow b_{1}, b_{2}$, and $\widehat{\rho}_{1}, \widehat{\rho}_{2}
\vdash \chi \Downarrow \widehat{\rho}'_{1}, \widehat{\rho}'_{2}; O$,
where $O$ are the observations (similar to source semantics). 
The former judgement evaluates a gate under the environments
$\widehat{\rho}_j$ and generates shares $b_j$ of the gate's output.
Rule ({\sc{C-In}}) evaluates the input gate $\kw{in}^{m}_{j}$, and
creates the $m$-type shares of the value $c$ input by party $j$.
Rule ({\sc{C-Mult}}) illustrates the pattern for evaluating circuit
gates $\gate$. To evaluate $\kw{mult}\:\gate_{1}\:\gate_{2}$, the
rule first evaluates $\gate_{1}$ to $(b_{11}, b_{21})$ and
$\gate_{2}$ to $(b_{12}, b_{22})$. Shares $(b_{11}, b_{21})$ are then
combined using $\mathcal{D_{\mathcal{A}}}$ to $n_{1}$, and similarly
$(b_{12}, b_{22})$ are combined to $n_{2}$. The final output of the
\kw{mult} gate is then $\mathcal{E}_{\mathcal{A}}(n_{1} \mulop n_{2})$. Note
that this is a functional description of how the \kw{mult} gate
evaluates, of course, concretely $n_{1}$ and $n_{2}$ are not observed
by the parties. Rule ({\sc{C-Coerce}}) creates the new shares using
$\mathcal{E}_m$ (the corresponding rule for coercion from
$\mathcal{P}$ is similar). The evaluation
of \kw{bind} updates the mapping of $\wire$ in the input environments, and
the rule ({\sc{C-Out}}) outputs the clear value $c$ to the
observations.

@@

The theorem states that if a source statement $s$ is well-typed, and
compiles to $\widetilde{s}$ in the intermediate language, then $s$ terminates in the source
semantics with observations $O$, $\widetilde{s}$ evaluates to circuit $\crct$, and $\crct$ terminates in the
circuit semantics with the same observations $O$. 
Therefore, well typed programs always terminate (array indices are always in bounds, there are no unbounded loops, etc.) and the \mpc protocol produces the same outputs as the source program.
% We first define a judgment for
%translation of environments, $\Gamma \vdash \rho \leadsto
%\widetilde{\rho}; \widehat{\rho}_{1}, \widehat{\rho}_{2}$. The
%judgment is defined inductively on $\Gamma$, and establishes that
%the static environment $\Gamma$ is consistent with the runtime
%environments $\rho$, $\widetilde{\rho}$, and $\widehat{\rho}_{1}$ and
%$\widehat{\rho}_{2}$. 
%As an example, if $x:\kw{uint}^{\mathcal{P}} \in
%\Gamma$, the judgment checks that $\rho(x) = \widetilde{\rho}(x) =
%n$, for some $n$. On the other hand, if $x:\kw{uint}^{m} \in
%\Gamma$ then the judgement checks that $\mathcal{D}_m( \widehat{\rho}_{1}(\widetilde{\rho}(x)),\widehat{\rho}_{2}(\widetilde{\rho}(x)))=\rho(x)$.
%%We elide the judgment, providing the full definition
%%in the supplementary material.
%With this judgment,
%the correctness theorem is as follows:
%
%\begin{theorem}[Correctness]\label{theorem:correctness}
%  $\forall\:\Gamma, s, \widetilde{s}, \Gamma_{1}, \rho, \widetilde{\rho}, \widehat{\rho}_{1}, \widehat{\rho}_{2}$, if:
%
%  \begin{enumerate}
%  \item $\Gamma \vdash s \leadsto \widetilde{s} \mid \Gamma_{1}$
%  \item $\Gamma \vdash \rho \leadsto \widetilde{\rho}; \widehat{\rho}_{1}, \widehat{\rho}_{2}$
%  \end{enumerate}
%
%  then $\exists \rho_{1}, O_{1}, \widetilde{\rho}_{1}, \chi, \widehat{\rho}'_{1}, \widehat{\rho}'_{2}, O_{2}$, s.t.
%
%  \begin{enumerate}[label=(\alph*)]
%  \item $\rho \vdash s \Downarrow \rho_{1}; O_{1}$
%  \item $\widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}_{1}; \chi$
%  \item $\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \chi \Downarrow \widehat{\rho}'_{1}, \widehat{\rho}'_{2}; O_{2}$
%  \item $\Gamma_{1} \vdash \rho_{1} \leadsto \widetilde{\rho}_{1}; \widehat{\rho}'_{1}, \widehat{\rho}'_{2}$
%  \item $O_{1} = O_{2}$
%  \end{enumerate}
%  
%\end{theorem}
%
%The theorem states that if a source statement $s$ is well-typed, and
%compiles to $\widetilde{s}$ in intermediate language, then $s$ terminates in the source
%semantics with observations $O_{1}$, $\widetilde{s}$ evaluates to circuit $\crct$, and $\crct$ terminates in the
%circuit semantics with observations $O_{2}$, where $O_{1} = O_{2}$
%%Furthermore, the final environments are also in the compilation
%%relation. \divya{Should we call the judgement between environs as compilation relation?}
%Due to space reasons, we provide the proof in \cite{anonymized}.

>>>>

\subsubsection*{Source language} Our source language
 is a simple imperative language shown in \figureref{srclang}. Types
$\psi$ consist of the base types $\sigma$, and arrays
of base types $\sigma[n]$, where $n$ is the array length. Though we
model only one dimensional arrays, our implementation supports
 multi-dimensional arrays as well. Expressions $e$ in the language include the
 integer constants $n$, \kw{bool} constants $\top$ and $\bot$,
variables $x$, binary operations $e_{1} \mulop e_{2}$ and $e_{1} >
e_{2}$ (we support several other
operators in the implementation, detailed in \sectionref{impl}),
conditionals $\cond{e}{e_{1}}{e_{2}}$,
array literals
$[\overline{e_{i}}]_{n}$\footnote{We write $\overline{e}$ (and
  similarly for other symbols) to denote a sequence of expressions.
The length of the sequence is usually clear from the context.}, and
array reads $x[e]$. The expression $\kw{in}_{j}$ denotes input from
party $j$. The statements $s$ in the language comprise of variable
declarations and assignments ($\psi\:x = e$ and $x := e$ resp.),
\kw{for} loops, array writes ($x[e_{1}] := e_{2}$), \kw{if}
statements, and sequence of statements ($s_{1}; s_{2}$). The statement
$\kw{out}\:e$ denotes revealing the value of $e$ to the
parties. The \kw{while} statement is an internal syntax that is not
exposed to the programmer. \\

@@

The judgment $\rho \vdash e \Downarrow v$ denotes the big-step
evaluation of an expression~$e$ to a value~$v$ under the runtime
environment~$\rho$. Rule ({\sc{E-Var}}) looks up the value of $x$ in
the environment. Rule ({\sc{E-Mult}}) inductively evaluates $e_{1}$ and
$e_{2}$, and returns their product. Rule ({\sc{E-Read}})
evaluates an array read operation. It first evaluates $x$ to an array
value $[\overline{c_{i}}]_{n_{1}}$, and $e$ to a \kw{uint} value
$n$. It then returns $c_{n}$, the $n$-th index value in the array,
provided $n < n_{1}$, the length of the array. Rule ({\sc{E-Inp}})
evaluates to some constant $c$ denoting party $j$'s input. 
%We model the inputs to be base constants, a
An array input can be written in the
language as $[\kw{in}_{j}]_{n}$, which can then evaluate using the
rule ({\sc{E-Arr}}) (the notation $\forall i \in [n]$ is read
  as $\forall i \in \{0\dots n - 1\}$). The remaining rules are
straightforward, and are elided for space reasons.

@@

The judgment $\Gamma \vdash e : \tau \leadsto \widetilde{e}$, where
$\Gamma$ maps variables $x$ to types $\tau$, says that
under $\Gamma$, $e$ (in the source language) compiles to $\widetilde{e}$
(in the intermediate language), where $\widetilde{e}$ has type $\tau$.
Rules ({\sc{T-UInt}}) and ({\sc{T-Bool}}) assigns the label
$\mathcal{P}$ to the constants, as the constants are always
public. Rule ({\sc{T-Mult}})
compiles a multiplication to either a public multiplication  ($\mulop_{\mathcal{P}}$), or a
secret arithmetic multiplication ($\mulop_{\mathcal{A}}$). As
our compiler is cryptographic cost aware, it never compiles the
multiplication to boolean multiplication $\mulop_{\mathcal{B}}$ (\sectionref{impl}). In a similar manner,
rule ({\sc{T-Gt}}) compiles $e_{1} > e_{2}$ to either public comparison,
or secret boolean comparison $>_\mathcal{B}$ (and never
$>_{\mathcal{A}}$).
%
The rule for conditional
({\sc{T-Cond}}) has two cases: when the conditional expression $e$ is of
type $\kw{bool}^{\mathcal{P}}$, both the branches have a base type
$\sigma^{\ell_{1}}$, for an arbitrary $\ell_{1}$, and the conditional is
compiled to a public conditional, whereas when the conditional
expression has type $\kw{bool}^{\mathcal{B}}$, $\ell_{1}$ is also
$\mathcal{B}$, and the conditional is compiled to a secret conditional
using a boolean circuit. Note that we restrict the type of the branches
to be of base type. 
Rule ({\sc{T-Read}}) type checks an array read. It checks
that the array index $e$ is public, and uses a static bounds checking
judgment $ \models e < n$ to prove that the array index is in
bounds. \sectionref{impl} discusses our implementation of this
check. Rule ({\sc{T-Inp}}) picks a label $m$ for the input. Finally,
the rule ({\sc{T-Sub}}) is the subsumption rule that coerces an expression of
type $\sigma^{\ell}$ to an expression of type $\sigma^{m}$ using the
coerce expression. It is important for security that the secrets
cannot be coerced to public values and indeed ({\sc{T-Sub}}) does not
permit it.

@@

We use $\widehat\rho_j$ to denote the circuit environment for party $j.$ 
We give the 
judgments $\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \gate
\Downarrow b_{1}, b_{2}$, and $\widehat{\rho}_{1}, \widehat{\rho}_{2}
\vdash \chi \Downarrow \widehat{\rho}'_{1}, \widehat{\rho}'_{2}; O$,
where $O$ are the observations (similar to source semantics). 
The former judgment evaluates a gate under the environments
$\widehat{\rho}_j$ and generates shares $b_j$ of the gate's output.
Rule ({\sc{C-In}}) evaluates the input gate $\kw{in}^{m}_{j}$, and
creates the $m$-type shares of the value $c$ input by party $j$.
Rule ({\sc{C-Mult}}) illustrates the pattern for evaluating circuit
gates $\gate$. To evaluate $\kw{mult}\:\gate_{1}\:\gate_{2}$, the
rule first evaluates $\gate_{1}$ to $(b_{11}, b_{21})$ and
$\gate_{2}$ to $(b_{12}, b_{22})$. Shares $(b_{11}, b_{21})$ are then
combined using $\mathcal{D_{\mathcal{A}}}$ to $n_{1}$, and similarly
$(b_{12}, b_{22})$ are combined to $n_{2}$. The final output of the
\kw{mult} gate is then $\mathcal{E}_{\mathcal{A}}(n_{1} \mulop n_{2})$. Note
that this is a functional description of how the \kw{mult} gate
evaluates, of course, concretely $n_{1}$ and $n_{2}$ are not observed
by the parties. Rule ({\sc{C-Coerce}}) creates the new shares using
$\mathcal{E}_m$ (the corresponding rule for coercion from
$\mathcal{P}$ is similar). The evaluation
of \kw{bind} updates the mapping of $\wire$ in the input environments, and
the rule ({\sc{C-Out}}) outputs the clear value $c$ to the
observations.

@@

The theorem states that if a source statement $s$ is well-typed, and
compiles to $\widetilde{s}$ in the intermediate language, then $s$ terminates in the source
semantics with observations $O$, $\widetilde{s}$ evaluates to circuit $\crct$, and $\crct$ terminates in the
circuit semantics with the same observations $O$. 
Therefore, well typed programs always terminate (array indices are always in bounds, there are no unbounded loops, etc.) and the \mpc protocol produces the same outputs as the source program.
% We first define a judgment for
%translation of environments, $\Gamma \vdash \rho \leadsto
%\widetilde{\rho}; \widehat{\rho}_{1}, \widehat{\rho}_{2}$. The
%judgment is defined inductively on $\Gamma$, and establishes that
%the static environment $\Gamma$ is consistent with the runtime
%environments $\rho$, $\widetilde{\rho}$, and $\widehat{\rho}_{1}$ and
%$\widehat{\rho}_{2}$. 
%As an example, if $x:\kw{uint}^{\mathcal{P}} \in
%\Gamma$, the judgment checks that $\rho(x) = \widetilde{\rho}(x) =
%n$, for some $n$. On the other hand, if $x:\kw{uint}^{m} \in
%\Gamma$ then the judgment checks that $\mathcal{D}_m( \widehat{\rho}_{1}(\widetilde{\rho}(x)),\widehat{\rho}_{2}(\widetilde{\rho}(x)))=\rho(x)$.
%%We elide the judgment, providing the full definition
%%in the supplementary material.
%With this judgment,
%the correctness theorem is as follows:
%
%\begin{theorem}[Correctness]\label{theorem:correctness}
%  $\forall\:\Gamma, s, \widetilde{s}, \Gamma_{1}, \rho, \widetilde{\rho}, \widehat{\rho}_{1}, \widehat{\rho}_{2}$, if:
%
%  \begin{enumerate}
%  \item $\Gamma \vdash s \leadsto \widetilde{s} \mid \Gamma_{1}$
%  \item $\Gamma \vdash \rho \leadsto \widetilde{\rho}; \widehat{\rho}_{1}, \widehat{\rho}_{2}$
%  \end{enumerate}
%
%  then $\exists \rho_{1}, O_{1}, \widetilde{\rho}_{1}, \chi, \widehat{\rho}'_{1}, \widehat{\rho}'_{2}, O_{2}$, s.t.
%
%  \begin{enumerate}[label=(\alph*)]
%  \item $\rho \vdash s \Downarrow \rho_{1}; O_{1}$
%  \item $\widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}_{1}; \chi$
%  \item $\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \chi \Downarrow \widehat{\rho}'_{1}, \widehat{\rho}'_{2}; O_{2}$
%  \item $\Gamma_{1} \vdash \rho_{1} \leadsto \widetilde{\rho}_{1}; \widehat{\rho}'_{1}, \widehat{\rho}'_{2}$
%  \item $O_{1} = O_{2}$
%  \end{enumerate}
%  
%\end{theorem}
%
%The theorem states that if a source statement $s$ is well-typed, and
%compiles to $\widetilde{s}$ in intermediate language, then $s$ terminates in the source
%semantics with observations $O_{1}$, $\widetilde{s}$ evaluates to circuit $\crct$, and $\crct$ terminates in the
%circuit semantics with observations $O_{2}$, where $O_{1} = O_{2}$
%%Furthermore, the final environments are also in the compilation
%%relation. \divya{Should we call the judgment between environs as compilation relation?}
%Due to space reasons, we provide the proof in \cite{anonymized}.
